Machine learning enhanced empirical potentials for metals and alloys

### 2020-04-22 doc/papers/ml-eam/manuscript.tex
The final version for npj.

\author{Xin Chen}
\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}

\author{Li-Fang Wang}
\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}

\author{Xing-Yu Gao}
\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}

\author{Ya-Fan Zhao}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}
\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}

\author{De-Ye Lin}
\email{lindeye0716@163.com}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}
\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}

\author{Wei-Dong Chu}
\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}

\author{Hai-Feng Song}
\email{song\_haifeng@iapcm.ac.cn}
\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Abstract
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{abstract}
Interatomic potential or force-field plays a vital role in simulating materials. 
Empirical potentials like the embedded atom method (EAM) and its variant 
angular-dependent potential (ADP) have proven successful in modeling many 
metals qualitatively. In the past few years, machine learning has became the 
key approach for modeling interatomic interactions. Powered by big data and 
efficient optimizers, machine learning interatomic potentials can generally 
approximate to the first-principles accuracy. In this works, we successfully 
developed a route to express EAM and ADP within machine learning framework in 
highly-vectorizable form and further enhanced the potentials with physical 
constraints. Machine learning can significantly boost their performances with 
few training data. 
For energy and force 
predictions, machine learned EAM and ADP can be almost as accurate as the 
spectral neighbor analysis potential (SNAP) on the fcc Ni, bcc Mo and Mo-Ni 
alloy systems, while the molecular dynamics simulations performed by the later 
approach is nearly 1000 times slower. Machine learned EAM and ADP can also give 
first-principle accuracy key material properties, such as elastic constants, 
melting temperatures and surface energies. Our results suggest a new and 
systematic route for developing machine learning interatomic potentials. 
\end{abstract}

@@

Suppose the cutoff radius $r_c$ is fixed, the energy of atom $i$ of specie $A$ 
can be calculated with the following expanded equation:
\begin{align}
\label{eq:e_atom_phase_1}
E_{i}^{A} 
= & 
\frac{1}{2}\sum_{j\neq i}^{\niaa}\phi_{\mathrm{AA}}(r_{ij}) +
\frac{1}{2}\sum_{j\neq i}^{\niab}\phi_{\mathrm{AB}}(r_{ij}) \nonumber \\
+ & F_{\mathrm{A}}\left(
    \sum_{j\neq i}^{\niaa}{\rho_{\mathrm{A}}(r_{ij})} +
    \sum_{j\neq i}^{\niab}{\rho_{\mathrm{B}}(r_{ij})} 
\right)
\end{align}
where $\niaa$ represents the number of A-type neighors of atom $i$ and 
$\niab$ represents the number of B-type neighors. For atom $j$ of specie $B$, we
can also write a similar form. 
\begin{align}
\label{eq:e_atom_phase_1}
E_{j}^{B} 
= & 
\frac{1}{2}\sum_{i \neq j}^{\njbb}\phi_{\mathrm{BB}}(r_{ij}) +
\frac{1}{2}\sum_{i \neq j}^{\njba}\phi_{\mathrm{AB}}(r_{ij}) \nonumber \\
+ & F_{\mathrm{B}}\left(
    \sum_{j\neq i}^{\njbb}{\rho_{\mathrm{B}}(r_{ij})} +
    \sum_{j\neq i}^{\njba}{\rho_{\mathrm{A}}(r_{ij})} 
\right)
\end{align}
When $r_c$ is fixed, $\niaa$, $\niab$, $\njbb$ and $\njba$ are all constants and
$\nnli$ is the maximum of these numbers of structure $i$. Hence, we can 
pre-determine the maximum neighor list size $\nnl$:
\begin{equation}
\label{eq:nnl}
\nnl = \max{\left(\nnli\right)}
\end{equation}
$\nnl$ is also a constant in the training phase because both the training 
dataset and $r_c$ are fixed.

@@

To conclude, we have successfully combined empirical potentials (EAM/ADP) with 
machine learning. The machine learning approaches (big data, SGD, etc), together 
with physical constraints (Rose EOS, elastic constants) can significantly 
improve the performaces of EAM/ADP. For the fcc Ni, bcc Mo and Mo-Ni alloys, 
ML-EAM and ML-ADP can be as accurate as SNAP, while later method is orders of 
magnitude slower. Our work also indicates a new route to design and develop 
machine learning interatomic potentials \textemdash the machine learning 
enhanced empirical potential approach.

@@

>>>>

\author{Xin Chen}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{Li-Fang Wang}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{Xing-Yu Gao}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{Ya-Fan Zhao}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{De-Ye Lin}
\email{lindeye0716@163.com}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{Wei-Dong Chu}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}

\author{Hai-Feng Song}
\email{song\_haifeng@iapcm.ac.cn}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Abstract
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{abstract}
Interatomic potential or force-field plays a vital role in simulating materials. 
Empirical potentials like the embedded atom method (EAM) and its variant 
angular-dependent potential (ADP) have proven successful in modeling many 
metals qualitatively. In the past few years, machine learning has became the 
key approach for modeling interatomic interactions. Powered by big data and 
efficient optimizers, machine learning interatomic potentials can generally 
approximate to the first-principle accuracy. In this works, we successfully 
developed a route to express EAM and ADP within machine learning framework in 
highly-vectorizable form and further enhanced the potentials with physical 
constraints. Machine learning can significantly boost their performances with 
few training data. 
For energy and force 
predictions, machine lear̦ned EAM and ADP can be almost as accurate as the 
spectral neighbor analysis potential (SNAP) on the fcc Ni, bcc Mo and Mo-Ni 
alloy systems, while the molecular dynamics simulations performed by our 
potentials are nearly 1000 times faster. Machine learned EAM and ADP can also 
give first-principle accuracy key material properties, such as elastic 
constants, melting temperatures and surface energies. Our results suggest a new 
and systematic route for developing machine learning interatomic potentials. 
\end{abstract}

@@

Suppose the cutoff radius $r_c$ is fixed, the energy of atom $i$ of specie $A$ 
can be calculated with the following expanded equation:
\begin{align}
\label{eq:e_atom_phase_1}
E_{i}^{A} 
= & 
\frac{1}{2}\sum_{j\neq i}^{\niaa}\phi_{\mathrm{AA}}(r_{ij}) +
\frac{1}{2}\sum_{j\neq i}^{\niab}\phi_{\mathrm{AB}}(r_{ij}) \nonumber \\
+ & F_{\mathrm{A}}\left(
    \sum_{j\neq i}^{\niaa}{\rho_{\mathrm{A}}(r_{ij})} +
    \sum_{j\neq i}^{\niab}{\rho_{\mathrm{B}}(r_{ij})} 
\right)
\end{align}
where $\niaa$ represents the number of A-type neighors of atom $i$ and 
$\niab$ represents the number of B-type neighors. For atom $j$ of specie $B$, we
can also write a similar form. 
\begin{align}
\label{eq:e_atom_phase_2}
E_{j}^{B} 
= & 
\frac{1}{2}\sum_{i \neq j}^{\njbb}\phi_{\mathrm{BB}}(r_{ij}) +
\frac{1}{2}\sum_{i \neq j}^{\njba}\phi_{\mathrm{AB}}(r_{ij}) \nonumber \\
+ & F_{\mathrm{B}}\left(
    \sum_{j\neq i}^{\njbb}{\rho_{\mathrm{B}}(r_{ij})} +
    \sum_{j\neq i}^{\njba}{\rho_{\mathrm{A}}(r_{ij})} 
\right)
\end{align}
When $r_c$ is fixed, $\niaa$, $\niab$, $\njbb$ and $\njba$ are all constants and
$\nnli$ is the maximum of these numbers of structure $i$. Hence, we can 
pre-determine the maximum neighor list size $\nnl$:
\begin{equation}
\label{eq:nnl}
\nnl = \max{\left(\nnli\right)}
\end{equation}
$\nnl$ is also a constant in the training phase because both the training 
dataset and $r_c$ are fixed.

@@

To conclude, we have successfully combined empirical potentials (EAM/ADP) with 
machine learning. The machine learning approaches (big data, optimization, etc), 
together with physical constraints (Rose EOS, elastic constants) can 
significantly improve the performaces of EAM/ADP. For the fcc Ni, bcc Mo and 
Mo-Ni alloys, ML-EAM and ML-ADP can be as accurate as SNAP, while later method 
is orders of magnitude slower. Our work also indicates a new route to design and 
develop machine learning interatomic potentials \textemdash the machine learning 
enhanced empirical potential approach.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Data Availability
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\section*{Data Availability}
The optimized parameters of the associated EAM and ADP models are all listed in 
the appendix. The learning datasets are originally created by Shyue Ping Ong 
and can be accessed from 
\href{https://github.com/materialsvirtuallab/snap}{GitHub} freely.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Author Contributions
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section*{Author Contributions}
X.C. designed the research project, wrote the related codes and build the 
machine learning interatomic potentials. L.F.W did the molecular dynamics 
simulations and calculations. X.Y.G., Y.F.Z., D.Y.L., W.D.C. and H.F.S. provided 
valuable comments and suggestions to the work. X.C. and L.F.W. wrote the paper. 
All authors contributed to discussions of the results.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Additional Information
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section*{Additional Information}

\textbf{Competing interests}: The authors declare no Competing Financial or 
Non-Financial Interests.
